MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-04-06 latex/paper/paper.tex
more carry

@@

    Following placement, the \textbf{routing} tool connects the placed elements by assigning physical wiring resources to the nets in the netlist. 
    The router determines their paths through the programmable interconnect network while attempting to meet timing and performance constraints.

@@

    \begin{figure*}[t]
        \centering
        \includegraphics[width=12.0cm]{figures/hierarchy.png}
        \caption{Architecture Hierarchy of a Xilinx FPGA (Ultrascale)}
        \label{fig:hierarchy}
    \end{figure*}

    Xilinx introduced the 7-Series architecture in 2010 which followed the heterogeneous island-style architecture described previously. 
    Then in 2013 Xilinx unveiled the Ultrascale architecture which also followed the heterogeneous island-style design but made several improvements over the 7-Series architecture. 
    Although the latest FPGAs follow the Ultrascale architecture, this paper focuses on the 7-Series FPGAs due to their greater accessibility for beginner users and compatibility with open source tooling.

    Figure~\ref{fig:hierarchy} illustrates the architectural hierarchy of an Ultrascale Xilinx FPGA. 
    The architecture hierarchy is the same for the 7-Series devices. 
    At its most basic level, an FPGA is a vast physical array of replicated atomic components known as Basic Elements of Logic (BELs). 
    These BELs, which include LUTs, FFs, BRAMs, DSPs, and programmable interconnects, form the fundamental building blocks for custom digital circuit implementations on FPGAs.

@@

    For a complete understanding of the RapidWright packages, please refer to the API Overview \url{https://www.rapidwright.io/docs/RapidWright_Overview.html}.

@@

        Furthermore, certain cell patterns are specific to certain cell types. 
        Therefore, we first traverse the entire EDIF netlist and identify all unique cell types and group them together via a {\tt HashMap<String, List<EDIFHierCellInst>>}, where String is the name cell type and {\tt EDIFHierCellInst} is the cell class.

        Cell groups can include {\tt CARRY4}, {\tt LUT1-6}, {\tt FF}, {\tt DSP48E1}, {\tt RAMB18E1}, and others. 
        We specifically look for CARRY4 chains, DSP48E1 cascades, and LUT-FF pairs as these patterns are common to nearly all FPGA designs.

@@

            As the name implies, CARRY chains are chains of CARRY cells. 
            Very often an HDL design will implement many adders, counters, subtractors, or comparators, all operations based on binary addition. 
            They are so ubiquitous that every that in the 7-Series architecture, every CLB comes embedded with a CARRY4 BEL, a 4-bit carry-lookahead (CLA) adder that can be chained together across CLBs to implement wider adders.
            On the Ultrascale architecture, the each CLB contains a CARRY8 BEL, an 8-bit CLA adder, allowing for even greater logic density.

@@

            These CARRY4 cells, if chained together, must be placed vertically across CLBs, as each CLB is connected vertically by Carry In (CI)
            This allows the facilitation of fast adders on the device.

@@

\section{Future Work}
    \begin{itemize}
    \item Different cooling schedules: linear cooling, logarithmic cooling, piecewise cooling.
    \item Support for more macros: buffer cells, FIFO, clock generators, etc..
    \item Support Ultrascale architecture. 
    \item Better packing scheme. Currently pseudorandom packing. 
    \item BEL-centric placement over Site-centric placement.
    \item Analytical placement. Global placement, legalization, detailed placement. 
    \end{itemize}

>>>>

    \begin{figure}[]
        \centering
        \includegraphics[width=9.0cm]{figures/hierarchy.png}
        \caption{Architecture Hierarchy of a Xilinx FPGA (Ultrascale)}
        \label{fig:hierarchy}
    \end{figure}

    Xilinx introduced the 7-Series architecture in 2010 which followed the heterogeneous island-style architecture described previously. 
    Then in 2013 Xilinx unveiled the Ultrascale architecture which also followed the heterogeneous island-style design and made several improvements over the 7-Series architecture. 
    Although the latest FPGAs follow the Ultrascale architecture, this paper focuses on the 7-Series FPGAs due to their greater accessibility for beginner users and compatibility with open source tooling.

    % Figure~\ref{fig:hierarchy} illustrates the architectural hierarchy of an Ultrascale Xilinx FPGA. 
    % The architecture hierarchy is the same for the 7-Series devices. 
    Figure \ref{fig:hierarchy_3} illustrates the architectural hierarchy of a 7-Series Xilinx FPGA. 
    At its most basic level, an FPGA is a vast physical array of replicated atomic components known as Basic Elements of Logic (BELs). 
    These BELs, which include LUTs, FFs, BRAMs, DSPs, and programmable interconnects, form the fundamental building blocks for custom digital circuit implementations on FPGAs.

    \begin{figure}[]
        \centering
        \includegraphics[width=9.0cm]{figures/hierarchy_3.png}
        \caption{Tiles, Sites, BELs hierarchy}
        \label{fig:hierarchy_3}
    \end{figure}

@@

    Following placement, the \textbf{routing} tool connects the placed elements by assigning physical wiring resources (switchboxes, programmable interconnects) to the nets in the netlist. 
    The router determines their paths through the programmable interconnect network while attempting to meet timing and performance constraints.

@@

@@

    For a complete understanding of the RapidWright packages, please refer to the API Overview \url{https://www.rapidwright.io/docs/RapidWright_Overview.html} and the API reference.

@@

        Furthermore, certain cell patterns are specific to certain cell types. 
        Therefore, we must first traverse the entire EDIF netlist and identify all unique cell types and group them together via a {\tt HashMap<String, List<EDIFHierCellInst>>}, where key String is the name cell type and value {\tt List<EDIFHierCellInst>} is the list of cells of that type. 
        The resulting lists in the hashmap are mutually exclusive.

        Cell types can include {\tt CARRY4}, {\tt LUT1-6}, {\tt FF}, {\tt DSP48E1}, {\tt RAMB18E1}, and others. 
        We specifically look for CARRY4 chains, DSP48E1 cascades, and LUT-FF pairs as these patterns are common to nearly all FPGA designs. 
        We will describe each cell pattern in detail.

@@

            As the name implies, CARRY chains are chains of CARRY cells. 
            Very often an HDL design will implement many adders, counters, subtractors, or comparators, all operations based on binary addition. 
            They are so ubiquitous that every that in the 7-Series architecture, every CLB comes embedded with a CARRY4 BEL, a 4-bit carry-lookahead (CLA) adder that can be chained together across CLBs to implement wider adders. 
            % On the Ultrascale architecture, the each CLB contains a CARRY8 BEL, an 8-bit CLA adder, allowing for even greater logic density.

@@

            These CARRY4 cells, if chained together, must be placed vertically across CLBs, as each CLB is connected vertically by a Carry In (CI) to Carry Out (COUT) wire between them. 
            The only way to route the carry net is through this wire, thus why these CLBs must be placed consecutively vertically. 
            This allows the facilitation of fast adders on the device.

@@

\section{Future Work}
    \begin{itemize}
    \item Different cooling schedules: linear cooling, logarithmic cooling, piecewise cooling.
    \item Support for more macros: buffer cells, FIFO, clock generators, etc..
    \item Support Ultrascale architecture. 
    \item Better packing scheme. Currently pseudorandom packing. 
    \item Inter-Site BEL swapping. 
    \item BEL-centric placement over Site-centric placement.
    \item Analytical placement. Global placement, legalization, detailed placement. 
    \end{itemize}
